Automatically drawing vegetation classification maps using digital time-lapse cameras in alpine ecosystems

### 2022-11-24 paper/paper.Rmd
modified paper

---
title: Automatically drawing vegetation classification maps using digital time-lapse cameras in alpine ecosystems
authors:
  - name: Ryotaro Okamoto
    thanks: https://github.com/0kam
    department: Doctoral Program in Biology
    affiliation: University of Tsukuba
    location: 1-1-1 Tennodai, Tsukuba, Ibaraki 305-8577 Japan. 
    email: okamoto.ryotaro.su@alumni.tsukuba.ac.jp
  - name: Hiroyuki Oguma
    department: Biodiversity Division
    affiliation: National Institute for Environmental Studies
    location: 16-2 Onogawa, Tsukuba, Ibaraki 305-8506 Japan
    email: oguma@nies.go.jp
  - name: Reiko Ide
    department: Earth System Division
    affiliation: National Institute for Environmental Studies
    location: 16-2 Onogawa, Tsukuba, Ibaraki 305-8506 Japan
    email: ide.reiko@nies.go.jp
abstract: |
  Alpine ecosystems are particularly vulnerable to the effect of climate change. Monitoring the distribution of alpine vegetation is required to plan practical conservation activities. However, conventional field observation and satellite remote sensing have difficulties in the coverage and frequency in alpine areas. Researchers have also utilized ground-based time-lapse cameras to observe alpine regions' snowmelt and vegetation phenology. Such time-lapse cameras have significant advantages in cost, resolution, and frequency. However, no research has used these to monitor the vegetation distribution. This study proposes a novel method to draw georeferenced vegetation maps from ground-based imagery of the Japanese alps. Our approach has two components: the classification and georectification method. The proposed classification method uses pixel time series acquired from autumn images to utilize the patterns of autumn foliage. We show that using time-lapse imagery and a Recurrent Neural Network improves the performance of the vegetation classification. We also developed a novel method to accurately transform a ground-based image into georeferenced data. We propose 1. an automated procedure to acquire Ground Control Points (GCPs) and 2. a camera model that considers the lens distortion for accurate georectification. We show that our approach outperforms the conventional method. The proposed method achieved sufficient accuracy to observe the vegetation distribution on a plant-community scale. The evaluation reveals an F1 score of 0.937 in the vegetation classification and a root mean square error (RMSE) of 3.4 m in the georectification. Our results highlight the potential of cheap time-lapse cameras to monitor the distribution of alpine vegetation. The proposed method should significantly contribute to effective conservation planning of the alpine ecosystems.
keywords:
  - alpine ecology
  - deep learning
  - ecosystem monitoring
bibliography: export.bib
biblio-style: unsrt
output:
  bookdown::pdf_book:
    fig_caption: yes
    base_format: rticles::arxiv_article

@@

```{r setup, include=FALSE}
library(tidyverse)
library(magrittr)
library(kableExtra)
library(tidyverse)
library(magrittr)
library(fs)
library(lubridate)
setwd("~/projects/VegetationMapPaper/paper/")
```

@@

The effects of climate change on terrestrial ecosystems are particularly significant in alpine regions (\cite{IPCC2007}). Alpine vegetation depends on severe conditions such as low temperatures and long snow-covered periods. Thus alpine areas have unique species adapted to such extreme environments. Several studies have reported that recent global climatic changes, e.g., increasing temperatures and reducing snow-covered periods, have accelerated the invasion of non-native species into alpine areas (see \cite{Alexander2016AlpBotany}). In Japan, dwarf bamboo (*Sasa kurilensis*) has invaded alpine snow meadows, probably driven by the extension of the snow-free period (\cite{Kudo2011EcoEvo}). Climate change has also affected the growth and phenologies of native species. For example, the growth of dwarf pine (*Pinus pumila*), a dominant species in Japanese alpine regions, has been affected by climatic conditions such as temperature and snowmelt (\cite{Amagai2015EcoRes}). In addition, researchers have reported that such changes vary depending on species and microhabitats (\cite{Kudo2010AAA}, \cite{XXXXXX}). Understanding and predicting the undergoing changes of alpine vegetation requires a monitoring method scalable to a wide range with a spatially and temporally high resolution.

@@

On the other hand, researchers have also utilized automated digital time-lapse cameras mounted on the ground for monitoring green-leaf phenologies in forests (\cite{Richardson2009EcolAppl}), grasslands (\cite{Browning2017RemSen}), and alpine meadows (\cite{IdeOguma2013EcolInfom}). Unlike satellite imagery, ground-based cameras provide images free of clouds and atmospheric effects. Also, they can obtain high-resolution (i.e., sub-meter scale) and frequent (i.e., daily or hourly) images at a meager cost. Most previous studies with such cameras were about calculating the phenology index (e.g., excess greenness, \cite{Woebbecke1995ASAE}) by setting some regions of interest (ROI) in the images. Few studies have utilized such images in monitoring vegetation distributions. This lack of studies seems to be because applying such time-lapse imagery in monitoring vegetation distribution has two technical challenges.

First, unlike satellites' multispectral sensors, ordinal digital cameras can only obtain three bands (Red, Green, and Blue), making it harder to classify the vegetation. Second, since digital time-lapse cameras are mounted on the ground, transforming these images into geospatial data (e.g., orthoimage, this process is called georectification) is challenging. In other words, even if we classify the vegetation from such images, we cannot quantitatively measure and analyze it using Geographic Information Systems (GIS) tools. Georectification is essential for utilizing time-lapse cameras in scientific conservation planning.

This study proposes an automated method for drawing vegetation classification maps with a digital time-lapse camera by solving these two challenges. We solved the first challenge by using time-lapse images to classify vegetation. The autumn leaf phenology, which varies among species, was utilized in the classification process. We show the effectiveness of such phenological information in vegetation classification. A novel georectification method for ground-based photographs was also developed to tackle the second challenge. Finally, the proposed method generated a vegetation classification map from time-lapse images taken in the Japanese alps. Our method enables us to continuously monitor vegetation distribution and its changes at a plant-community scale, covering broad areas at a meager cost.

@@

Our approach has three steps:
Image-to-image alignment for correcting the movements of the camera
Vegetation classification for classifying each pixel into vegetation categories
Georectification for transforming vegetation classification results into geospatial data
Fig. \ref{fig:diagram} shows the overview of the entire process. The proposed method requires training data for the vegetation classification model, a Digital Elevation Model (DEM), and a georectified airborne/satellite image that covers the camera's field of view.

@@

Our method is intended for digital time-lapse images of alpine landscapes. In this study, we used time-lapse photos of a Japanese alpine region owned by the National Institute for Environmental Studies, Japan (NIES). All the images are publically available on NIES' webpage (https://db.cger.nies.go.jp/gem/ja/mountain/station.html?id=2). In 2010, NIES installed the digital time-lapse camera (EOS 5D MK2, Canon Inc., 21 M pixels) on a mountain lodge Murodo-sanso (about 2350 m a.s.l., above the forest limit), located at the foot of Mt. Tateyama (3015 m a.s.l.), in the Nothern Japanese Alps (Fig. \ref{fig:map}). The photographs were taken hourly, from 6 a.m. to 7 p.m. The camera's field of view (FOV) includes Mt. Tateyama, which ranges from about 2350 m a.s.l. to 3015 m a.s.l. in elevation. The pixel size on the ground was about 0.5 m at the ridge, about 1 km from the cameras. The area has a complex mosaic-like vegetation structure because of its topography, including rocks, cliffs, curls, and moraines. From April to November, the camera observed the snowmelt and seasonal phenology of evergreen(e.g., *Pinus pumila*) and deciduous (*Sorbus* sp.,  *Betula ermanii*) dwarf trees, dwarf bamboos (e.g., *Sasa kurilensis*), and alpine shrubs and herbaceous plants (e.g., *Geum pentapetalum*, *Nephrophyllidium crista-galli*).

@@

### Image selection
We selected seven images from late summer to late fall (8/25, 9/5, 9/12, 9/20, 9/26, 10/3, and 10/10) in 2015. Since the temporal patterns of autumn foliage coloration vary among species (see Fig. \ref{fig:pixtimeseries}), this season is suitable for classifying vegetation. While acquiring a series of images costs a lot with UAVs and airborne imagery, time-lapse cameras are particularly good at observing such temporal patterns of leaf color.

@@

Although the camera is fixed on the mountain lodge, the images are slightly misalighned do to the wind. To handle this issue, we aligned the images before processing using Python3 programming language and OpenCV4 (https://opencv.org/) image processing library. First, one image of 2015 were set as the alignment target. Next, we automatically matched key points between the target and other images using the AKAZE local feature extractor (\cite{Alcantarilla2013AKAZE}) and K-nearest neighbor matcher. Then, we searched and applied the homography matrix that minimizes the distance between each pair of the matching points, using OpenCV's "findHomography" function. Applying the estimated homography matrix, we could align the images accurately (0.654 pixels in root mean square error (RMSE) of the matching points). Finally, an input mask were applied to define image regions that should be ignored in the following procedure, such as the sky and the regions too close to the camera. This process ensures that a set of pixel time-series reflects a phenological information of a same location, which is essential in the vegetation classification process.

## Automatic vegetation classification

Because the images were aligned succesfuly, we can stack the images and extract a time series of pixel values (Red, Green, and Blue, Fig. \ref{fig:pixtimeseries}) for each pixel. Such pixel time series reflects the temporal patterns of leaf colors. Because deciduous plants have great differences in autumn leaf phenology among species, researchers have used that information for vegetation classification with satellite imagery (\cite{Tigges2013RemSenEnv}, \cite{Son2013RemSen}, \cite{Heupel2018PFG}). However, no research has applied this technique to ground-based time-lapse imagery. We implemented Support Vector Machine (SVM) and Recurrent Neural Network (RNN) based vegetation classifiers, and tested the effects of using pixel time series of the ground-based imagery on the classification performance.

(ref:pixtimeseries) Pixel time series acquired from the time-lapse camera. You can obtain a time series of pixel values (i.e., Red, Green, and Blue) for each pixel on the photographs. Such pixel time series reflects the autumn phenology of the vegetation.

```{r, pixtimeseries, echo=FALSE, message=FALSE, fig.cap="(ref:pixtimeseries)", out.width="100%", auto_pdf = TRUE}
knitr::include_graphics("paper_files/figures/Slide1.jpg")
```

@@

We prepared two supervised models, SVM and RNN, to classify the pixel time series into vegetation categories. SVM is one of the most popular machine learning models in remote sensing and has many applications (\cite{Mountrakis2011SVMReview}), including vegetation classification with multitemporal satellite imagery (\cite{Tigges2013RemSenEnv}). RNN is a neural network that recognizes temporal or sequential data dynamics. Researchers have also utilized RNN for remote sensing tasks, such as land cover classification, with multi-temporal satellite imagery (\cite{Ienco2017RemSenLSTM}, \cite{Sharma2018NN}). \cite{Ienco2017RemSenLSTM} reported that an RNN classifier outperforms an SVM classifier. Amongst many variants of RNN, we used Long Short Time Memory (LSTM, \cite{Hochreiter1997LSTM}), one of the most well-known RNN architectures. Batch-normalization technique (\cite{IoffeSzegedy2015BatchNorm}) was employed to speed up and stabilize the model training. We also classified the pixel of every single image separately using SVM classifiers to test whether using multi-temporal imagery improves the classification performance.

@@

We set 7 vegetation classes: Dwarf Pine (*Pinus pumila*), Dwarf Bamboo (*Sasa kurilensis*), Rowans (*Sorbus sambucifolia*, *Sorbus matsumurana*), Birch (*Betula ermanii*), Alder (*Alnus viridis* subsp. *maximowiczii*), Other Vegetations (such as alpine shrubs and herbaceous plants), and No Vegetation. Using an open-source image annotation software (Semantic Segmentation Editor, Hitachi, https://github.com/Hitachi-Automotive-And-Industry-Lab/semantic-segmentation-editor), an expert prepared a teacher dataset for each class. Then we validated the teacher dataset with a set of telephotos taken in the summer and fall of 2016.

### Implementation and model training

We implemented the classifier with Python3 language, ThunderSVM library (<https://github.com/Xtra-Computing/thundersvm>) for the SVM classifier, and  PyTorch deep neural network library (<https://pytorch.org/>) for the RNN classifier. All source codes are publically available via GitHub (<https://github.com/0kam/VegetationMapPaper>).

@@

Then, we developed a novel method to convert ground-based landscape imagery into geographical data. This process is called georectification. Georectification of ground-based images has been a difficult task, and this causes the underuse of potentially rich information in ground-based imagery. In plain words, georectification means aligning images onto Digital Surface Models (DSMs) so that every pixel of an image gets a geographical coordinate. You can do this by modeling the camera. A camera is considered a function that transforms 3D geographical coordinates (e.g., X, Y, and height in a Universal Transverse Mercator coordinate system (UTM)) into 2D image coordinates (locations of each pixel in the image). Estimating the parameters of this function (such as the camera location, pose, and the FoV), you can simulate how each point of the DSM appears in the image. Usually, georectification has three steps:

1.  Finding Ground Control Points (GCPs) in the image.\
2.  Estimating the camera parameters such as camera poses and FoV using GCPs.\
3.  Projecting the DSM into the image using the camera parameters.

Recently, researchers have developed some georectification methods to use ground-based photographs in glaciology (\cite{Messerli2015GeoInst}) and snow cover studies (\cite{Portenier2020Cryosphere}). Especially, \cite{Portenier2020Cryosphere} is worth mentioning for its semi-automatic method using mountain silhouettes as GCPs since setting GCPs in alpine region is not easy. However, this silhouette-based method has a drawback in the projection accuracy. It only uses limited areas (silhouettes) of images in the GCP acquisition, and also it ignores lens distortion. Because our target site has a complex vegetation distribution and our camera has considerable lens distortion, we needed a more accurate method.

@@

As we mentioned before, we modeled a camera as a function that transforms the geographical coordinates of the target mountain into the image coordinates in the acquired image. We implemented this procedure using the OpenGL framework (<https://www.opengl.org/>) to accelerate it with graphical processing units (GPUs). In OpenGL, we can divide this process into three operations:

1. Transforming the world (geographical) coordinates to the view coordinates (coordinates seen from the camera's point of view) using the camera's extrinsic parameters (i.e., the location and angles of the camera). \
2. Distorting the view coordinates using the lens distortion parameters. \
3. Transforming the view coordinates to the screen (image) coordinates using the camera's intrinsic parameters (i.e., the FoV of the camera).

First, we transformed the absolute geographic coordinates (also known as the world coordinates) into the view coordinates relative to the camera's position and direction. We applied a $4 \times 4$ view matrix $M_{view}$ (\ref{view_matrix}) to the geographic coordinates in this step. The view matrix represents the camera's position and direction (pan, tilt, roll). The geographic coordinate system must be cartesian (such as the UTM coordinate systems).

@@

Where $pan, tilt, roll$ are the Euler angles of the camera pose and $x, y, z$ are the camera location in the geographic coordinate system. Then, we can transform the geographic coordinates of the DSM $\begin{bmatrix} X_{geo} & Z_{geo} & Y_{geo} & 1 \end{bmatrix}$ to the view coordinates $\begin{bmatrix} X_{view} & Z_{view} & Y_{view} & 1 \end{bmatrix}$ applying the view matrix $M_{view}$ (\ref{view_tf}). In the OpenGL's view coordinate system, $X_{view}$, $Z_{view}$ and $Y_{view}$ represents horizontal, vertical, and depth positions respectively.

@@

Second, we distorted the camera coordinates to simulate the lens distortion. We modeled the lens distortion (\ref{dist_model}) based on \cite{Weng1992CameraCalib} and OpenCV's implementation (<https://docs.opencv.org/4.x/d9/d0c/group__calib3d.html>), where $X_{norm}$ and $Z_{norm}$ are the $Y$-normalized view coordinates. This model distorts points' locations according to the distance from the center of the image when projected to the image surface. Our model includes radial ($k1$~$k6$), tangental ($p1$, $p2$), thin prism ($s1$~$s4$) distortion, and unequal pixel aspect ratio ($a1$, $a2$). See \cite{Weng1992CameraCalib} for the details of lens distortion modeling.

>>>>

---
title: Automatically drawing vegetation classification maps using digital time-lapse cameras in alpine ecosystems
authors:
  - name: Ryotaro Okamoto
    thanks: https://github.com/0kam
    department: Doctoral Program in Biology
    affiliation: University of Tsukuba
    location: 1-1-1 Tennodai, Tsukuba, Ibaraki 305-8577 Japan. 
    email: okamoto.ryotaro.su@alumni.tsukuba.ac.jp
  - name: Hiroyuki Oguma
    department: Biodiversity Division
    affiliation: National Institute for Environmental Studies
    location: 16-2 Onogawa, Tsukuba, Ibaraki 305-8506 Japan
    email: oguma@nies.go.jp
  - name: Reiko Ide
    department: Earth System Division
    affiliation: National Institute for Environmental Studies
    location: 16-2 Onogawa, Tsukuba, Ibaraki 305-8506 Japan
    email: ide.reiko@nies.go.jp
abstract: |
 Alpine ecosystems are particularly vulnerable to the effect of climate change. Monitoring the distribution of alpine vegetation is required to plan practical conservation activities. However, conventional field observation and satellite remote sensing have difficulties in the coverage and frequency in alpine areas. Researchers have also utilized ground-based time-lapse cameras to observe alpine regions' snowmelt and vegetation phenology. Such time-lapse cameras have significant advantages in cost, resolution, and frequency. However, no research has used these to monitor vegetation distribution. This study proposes a novel method to draw georeferenced vegetation maps from ground-based imagery of the Japanese alps. Our approach has two components:classification and georectification methods. The proposed classification method uses pixel time series acquired from autumn images to utilize the patterns of autumn foliage. We show that using time-lapse imagery and a Recurrent Neural Network improves the performance of the vegetation classification. We also developed a novel method to transform a ground-based image into georeferenced data accurately. We propose 1. an automated procedure to acquire Ground Control Points (GCPs) and 2. a camera model that considers the lens distortion for accurate georectification. We show that our approach outperforms the conventional method. The proposed method achieved sufficient accuracy to observe the vegetation distribution on a plant-community scale. The evaluation reveals an F1 score of 0.937 in the vegetation classification and a root-mean-square error (RMSE) of 3.4 m in the georectification. Our results highlight the potential of cheap time-lapse cameras to monitor the distribution of alpine vegetation. The proposed method should significantly contribute to effective conservation planning of the alpine ecosystems.
keywords:
  - alpine ecology
  - deep learning
  - ecosystem monitoring
bibliography: export.bib
biblio-style: unsrt
output:
  bookdown::pdf_book:
    fig_caption: yes
    base_format: rticles::arxiv_article

@@

```{r setup, include=FALSE}
library(tidyverse)
library(magrittr)
library(kableExtra)
library(tidyverse)
library(magrittr)
library(fs)
library(lubridate)
setwd("~/Projects/VegetationMapPaper/paper/")
```

@@

The effects of climate change on terrestrial ecosystems are particularly significant in alpine regions (\cite{IPCC2007}). Alpine vegetation depends on severe conditions such as low temperatures and long snow-covered periods. Thus alpine areas have unique species adapted to such extreme environments. Several studies have reported that recent global climatic changes, e.g., increasing temperatures and reducing snow-covered periods, have accelerated the invasion of non-native species into alpine areas (see \cite{Alexander2016AlpBotany}). In Japan, dwarf bamboo (*Sasa kurilensis*) has invaded alpine snow meadows, probably driven by the extension of the snow-free period (\cite{Kudo2011EcoEvo}). Climate change has also affected the growth and phenologies of native species. For example, the growth of dwarf pine (*Pinus pumila*), a dominant species in Japanese alpine regions, have been affected by climatic conditions such as temperature and snowmelt (\cite{Amagai2015EcoRes}). In addition, researchers have reported that such changes vary depending on species and microhabitats (\cite{Kudo2010AAA}). Understanding and predicting the undergoing changes of alpine vegetation requires a monitoring method scalable to a wide range with a spatially and temporally high resolution.

@@

On the other hand, researchers have also utilized automated digital time-lapse cameras mounted on the ground for monitoring green-leaf phenologies in forests (\cite{Richardson2009EcolAppl}), grasslands (\cite{Browning2017RemSen}), and alpine meadows (\cite{IdeOguma2013EcolInfom}). Unlike satellite imagery, ground-based cameras provide images free of clouds and atmospheric effects. Also, they can obtain high-resolution (i.e., sub-meter scale) and frequent (i.e., daily or hourly) images at a meager cost. Most previous studies with such cameras were about calculating the phenology index (e.g., excess greenness, \cite{Woebbecke1995ASAE}) by setting some regions of interest (ROI) in the images. Few studies have utilized such images in monitoring vegetation distributions. This lack of studies seems to be because applying such time-lapse imagery in monitoring vegetation distribution has two technical challenges. First, unlike satellites' multispectral sensors, ordinal digital cameras can only obtain three bands (Red, Green, and Blue), making it harder to classify the vegetation. Second, since digital time-lapse cameras are mounted on the ground, transforming these images into geospatial data (e.g., orthoimage, this process is called georectification) is challenging. In other words, even if we classify the vegetation from such images, we cannot quantitatively measure and analyze it using Geographic Information Systems (GIS) tools. Georectification is essential for utilizing time-lapse cameras in scientific conservation planning.

This study proposes an automated method for drawing vegetation classification maps with a digital time-lapse camera by solving these two challenges. We solved the first challenge by using time-lapse images to classify vegetation. The autumn leaf phenology, which varies among species, was utilized in the classification process. We show the effectiveness of such phenological information in vegetation classification. A novel georectification method for ground-based photographs was also developed to tackle the second challenge. Using the proposed method, we generated a vegetation classification map from time-lapse images taken in the Japanese alps. Our approach enables us to continuously monitor vegetation distribution and its changes at a plant-community scale, covering broad areas at a meager cost.

@@

The proposed method has three steps:
1. Image-to-image alignment for correcting the movements of the camera.
2. Vegetation classification for classifying each pixel into vegetation categories.
3. Georectification for transforming vegetation classification results into geospatial data.
Fig. \ref{fig:diagram} shows the overview of the entire process. The proposed method requires training data for the vegetation classification model, a Digital Elevation Model (DEM), and a georectified airborne/satellite image that covers the camera's field of view.

Fig. \ref{fig:diagram} shows the overview of the entire process. The proposed method requires a teacher dataset for training the vegetation classification model, a Digital Elevation Model (DEM), and a georectified airborne/satellite image that covers the camera's field of view for the georectification process.

@@

Our approach is intended for digital time-lapse images of alpine landscapes. In this study, we demonstrated it with time-lapse photos of a Japanese alpine region owned by the National Institute for Environmental Studies, Japan (NIES). All the images are publically available on NIES' webpage (https://db.cger.nies.go.jp/gem/ja/mountain/station.html?id=2). In 2010, NIES installed the digital time-lapse camera (EOS 5D MK2, Canon Inc., 21 M pixels) on a mountain lodge Murodo-sanso (about 2350 m a.s.l., above the forest limit), located at the foot of Mt. Tateyama (3015 m a.s.l.), in the Nothern Japanese Alps (Fig. \ref{fig:map}). The photographs were taken hourly, from 6 a.m. to 7 p.m. The camera's field of view (FoV) includes Mt. Tateyama, which ranges from about 2350 m a.s.l. to 3015 m a.s.l. in elevation. The pixel size on the ground was about 0.5 m at the ridge, about 1 km from the cameras. The area has a complex mosaic-like vegetation structure because of its topography, including rocks, cliffs, curls, and moraines. From April to November, the camera observed the snowmelt and seasonal phenology of evergreen(e.g., *Pinus pumila*) and deciduous (*Sorbus* sp., *Betula ermanii*) dwarf trees, dwarf bamboos (e.g., *Sasa kurilensis*), and alpine shrubs and herbaceous plants (e.g., *Geum pentapetalum*, *Nephrophyllidium crista-galli*).

@@

### Image selection

In the beginning, we selected seven images with good weather from late summer to late fall (25th Aug, 5th, 12th, 20th, 26th Sep, 3rd, and 10th Oct) in 2015. Since the temporal patterns of autumn foliage coloration vary among species (Fig. \ref{fig:pixtimeseries}), this season is suitable for classifying vegetation. While acquiring a series of images costs a lot with UAVs and airborne imagery, time-lapse cameras are particularly good at observing such temporal patterns of leaf color.

(ref:pixtimeseries) Pixel time series acquired from the time-lapse camera. You can obtain a time series of pixel values (i.e., Red, Green, and Blue) for each pixel on the photographs (left). Such pixel time series reflects the autumn phenology of the vegetation and varies among vegetation classes (right).
```{r, pixtimeseries, echo=FALSE, message=FALSE, fig.cap="(ref:pixtimeseries)", out.width="100%", auto_pdf = TRUE}
knitr::include_graphics("paper_files/figures/Slide1.jpg")
```

@@

Time-lapse images often suffer from the camera's movement caused by strong wind, human interference, and thermal expansion of the camera platform. Since such movement can cause errors in vegetation classification and georectification, we implemented an alignment method that aligns each image to one image called the master image. The software was developed using Python3 programming language and OpenCV4 (https://opencv.org/) image processing library. We employed a feature-based alignment method used by previous studies (\cite{Hulton2020PyTrx}, \cite{Portenier2020Cryosphere}). First, we searched and matched key points between an image and the master image using the AKAZE local feature extractor (\cite{Alcantarilla2013AKAZE}) and K-nearest neighbor matcher. Then, we searched and applied the homography matrix that minimizes the distance between each pair of the matching points, using OpenCV's "findHomography" function (\ref{homography}). Applying the estimated homography matrix, we could align the images accurately (0.654 pixels in root mean square error (RMSE) of the matching points). Finally, an input mask was applied to define image regions that should be ignored in the following procedure, such as the sky and the regions too close to the camera. This process ensures that a set of pixel time series reflects phenological information of the same location, which is essential in the vegetation classification process.

(ref:homography) A homography matrix. $u$ and $v$ indicates the x and y-axis coordinates of a point on an image. We optimized $H$ by minimizing the Euclid distance between the transformed coordinate of a key point $\begin{bmatrix}{u', v'}\end{bmatrix}$ and the corresponding key point on the master image.

```{=tex}
\label{homography}
\begin{equation}
  \begin{bmatrix} 
    u' \\ v' \\ w'
  \end{bmatrix}
  =
  H
  \begin{bmatrix} 
    u \\ v \\ w 
  \end{bmatrix}
\end{equation}
```

## Vegetation classification

After the image-to-image alignment, we stuck the images and extract a time series of pixel values (Red, Green, and Blue, see Fig. \ref{fig:pixtimeseries}) for each pixel. Such pixel time series reflects the temporal patterns of leaf colors. Because deciduous plants have great differences in autumn leaf phenology among species, researchers have used that information for vegetation classification with satellite imagery (\cite{Son2013RemSen}, \cite{Tigges2013RemSenEnv}, \cite{Heupel2018PFG}). However, no research has applied this technique to ground-based time-lapse imagery. We implemented Support Vector Machine (SVM) and Recurrent Neural Network (RNN) based vegetation classifiers and tested the effects of using pixel time series of the ground-based imagery on the classification performance.

@@

We prepared two supervised models, SVM and RNN, to classify the pixel time series into vegetation categories. SVM is one of the most popular machine-learning models in remote sensing and has many applications (\cite{Mountrakis2011SVMReview}), including vegetation classification with multitemporal satellite imagery (\cite{Tigges2013RemSenEnv}). RNN is a neural network that recurrently processes sequential or temporal steps of data to recognize their dynamics. RNN has been used in many tasks such as speech recognition (\cite{Graves2013SpeechRNN}, \cite{Graves2014SpeechRNN}, \cite{Sak2014acousticLSTM}), and machine translation (\cite{Auli2013translationRNN}, \cite{Cho2014RNN}). Researchers have also utilized RNN for remote sensing tasks, such as land cover classification, with multi-temporal satellite imagery (\cite{Ienco2017RemSenLSTM}, \cite{Sharma2018NN}). (\cite{Ienco2017RemSenLSTM}) reported that an RNN classifier outperforms an SVM classifier on the land cover classification accuracy. Amongst many variants of RNN, we used Long Short-Time Memory (LSTM, \cite{Hochreiter1997LSTM}, one of the most well-known RNN architectures. Batch-normalization technique (\cite{IoffeSzegedy2015BatchNorm}) and the GELU activation function (\cite{HendrycksGimpel2016GELU}) were employed to speed up and stabilize the model training. As a comparison, we also classified the pixel of every single image separately using SVM classifiers to test whether using multi-temporal imagery improves the classification performance.

@@

Each pixel time series was classified into 7 vegetation classes: Dwarf Pine (*Pinus pumila*), Dwarf Bamboo (*Sasa kurilensis*), Rowans (*Sorbus sambucifolia*, *Sorbus matsumurana*), Maple (*Acer tschonoskii*), Montane Alder (*Alnus viridis* subsp. *maximowiczii*), Other Vegetations (such as alpine shrubs and herbaceous plants), and No Vegetation. Experts prepared a teacher dataset for each class using an open-source image annotation software (Semantic Segmentation Editor, Hitachi, https://github.com/Hitachi-Automotive-And-Industry-Lab/semantic-segmentation-editor). The prepared teacher dataset contains XX pixels in XX polygons, which covers about X% of the image. Each polygon was validated with telephotos taken at the same place in 2016. For some polygons that were not seen clearly even in the telephoto, we conducted a field survey in the November of 2022.

### Experimental Settings
We implemented the classifiers with Python3 language. For the SVM classifier, we used the ThunderSVM library (<https://github.com/Xtra-Computing/thundersvm>) and an RBF kernel with a regularization parameter equal to 1.0. The gamma value was set automatically by the “scaled“ setting of the ThunderSVM library as $\gamma = \frac{1}{n_{features} \times V(X)}$, where and stand for the dimensions and variance of the input data respectively. The RNN classifier was developed using the PyTorch deep neural network library (<https://pytorch.org/>). We set the number of hidden dimensions equal to 5. We trained the model 200 epochs with the RAdam optimizer (\cite{Liu2020RAdam}) with a learning rate of 0.001 and a batch size of 500. To validate the performance of the different classifiers, we performed a 5-fold cross-validation. Each fold was stratified with the vegetation categories. As a model evaluation metric, we used the per-class F1-score, a standard metric in machine learning (\ref{f1score}). $TP, FP, FN$ stand for the numbers of true positives, false positives, and false negatives, respectively. All source codes are available via GitHub (https://github.com/0kam/VegetationMapPaper).

```{=tex}
\label{f1score}
\begin{align}
\label{metrics}
  \begin{gathered}
  precision = \frac{TP}{TP + FP} \\
  recall = \frac{TP}{TP + FN} \\
  F_1 = \frac{2 \cdot precision \cdot recall}{precision + recall} \\
  \end{gathered}
\end{align}
```

@@

The second challenge of utilizing time-lapse cameras in vegetation monitoring is the difficulty of georectification, i.e., georeferencing and transforming the images into geographic data. We developed a novel method to perform this process automatically and accurately. Georectification of ground-based images has been a difficult task, and this causes the underuse of potentially rich information in ground-based imagery. In plain words, georectification means aligning images onto Digital Surface Models (DSMs) so that every pixel of an image gets a geographical coordinate. It usually requires a camera model. A camera is considered a function that transforms 3D geographical coordinates (e.g., X, Y, and height in a Universal Transverse Mercator coordinate system (UTM)) into 2D image coordinates (locations of each pixel in the image). Estimating the parameters of this function (such as the camera location, pose, and the FoV), you can simulate how each point of the DSM appears in the image. Usually, georectification has three steps:

1. Finding Ground Control Points (GCPs) in the image.
2. Estimating the camera parameters such as the camera pose and FoV using GCPs.
3. Projecting the image onto the DSM using the camera parameters.

Recently, researchers have developed some georectification methods to use ground-based photographs in glaciology (\cite{Messerli2015GeoInst}, \cite{Hulton2020PyTrx}) and snow cover studies (\cite{Portenier2020Cryosphere}). Usually, GNSS-positioned GCPs (such as GCP markers) are set before photo acquisition to acquire GCPs. However, setting such markers in alpine areas is often difficult due to their rugged topology. (\cite{Portenier2020Cryosphere}) tackled this problem by developing a method that uses mountain silhouettes to match images and DSMs directly. However, this silhouette-based method has a drawback in projection accuracy. It ignores lens distortion and only uses limited areas (silhouettes) of images in the GCP acquisition. To overcome these challenges, This study proposes 1. a camera model which includes lens distortion and 2. a novel method to acquire GCPs in a broader area than the silhouette-based method.

@@

In this section, we explain our camera model with lens distortion modeling. As mentioned, we modeled a camera as a function that transforms geographical coordinates into image coordinates. We implemented this procedure using the OpenGL framework (https://www.opengl.org) and the ModernGL Python library (\cite{moderngl}) to accelerate it with graphical processing units (GPUs). In OpenGL, we can divide this process into three operations:

1. Transforming the geographical coordinates (also called the world coordinates) to the view coordinates (coordinates seen from the camera's point of view) using the camera's extrinsic parameters (i.e., the location and angles of the camera).
2. Distorting the view coordinates using the lens distortion parameters.
3. Transforming the view coordinates to the image coordinates (also called the screen coordinates) using the camera's intrinsic parameters (i.e., the camera's FoV).

First, we transformed the geographic coordinates into the view coordinates, which are relative to the camera's position and direction. We used a $4 \times 4$ view matrix (\ref{view_matrix}) $M_{view}$ in this step. The view matrix represents the camera's position and direction (pan, tilt, roll).

@@

Where $pan, tilt, roll$ are the Euler angles of the camera pose and $x, y, z$ are the camera location in the geographic coordinate system. Applying the view matrix $M_{view}$ transforms the geographic coordinates  $\begin{bmatrix} X_{geo} & Z_{geo} & Y_{geo} & 1 \end{bmatrix}$ (the horizontal, vertical, and depth coordinates, respectively) into the view coordinates $\begin{bmatrix} X_{view} & Z_{view} & Y_{view} & 1 \end{bmatrix}$ (\ref{view_tf}). In the OpenGL's view coordinate system, $X_{view}$, $Z_{view}$ and $Y_{view}$ represents horizontal, vertical, and depth positions respectively. Note that the geographic coordinate system must be cartesian (such as the UTM coordinate systems, not the lat/long coordinate system).

@@

Second, we distorted the view coordinates to simulate the lens distortion. We modeled the lens distortion (\ref{dist_model}) based on \cite{Weng1992CameraCalib} and OpenCV's implementation (<https://docs.opencv.org/4.x/d9/d0c/group__calib3d.html>), where $X_{norm}$ and $Z_{norm}$ are the $Y$-normalized view coordinates. This model distorts points' locations according to the distance from the center of the image when projected to the image surface. Our model includes radial ($k1$~$k6$), tangental ($p1$, $p2$), thin prism ($s1$~$s4$) distortion, and unequal pixel aspect ratio ($a1$, $a2$). See \cite{Weng1992CameraCalib} for the details of lens distortion modeling.

```{=tex}
\begin{gather}
\label{dist_model}
  \begin{gathered}
  X_{norm} = \frac{X_{view}}{Y_{view}} \\
  Z_{norm} = \frac{Z_{view}}{Y_{view}} \\
  r^2 = {X_{norm}}^2 + {Z_{norm}}^2 \\
  \begin{bmatrix}
    X_{dist\_norm} \\ 
    Z_{dist\_norm} \\
  \end{bmatrix} 
  = 
  \begin{bmatrix} 
    X_{norm} \frac{1 + k_1 r^2 + k_2 r^4 + k_3 r^6}{1 + k_4 r^2 + k_5 r^4 + k_6 r^6} + 2 p_1 x’ y’ + p_2(r^2 + 2 x’^2) + s_1 r^2 + s_2 r^4 \\ 
    Z_{norm} \frac{1 + a_1 + k_1 r^2 + k_2 r^4 + k_3 r^6}{1 + a_2 + k_4 r^2 + k_5 r^4 + k_6 r^6} + p_1 (r^2 + 2 y’^2) + 2 p_2 x’ y’ + s_3 r^2 + s_4 r^4 \\    \end{bmatrix} \\
  X_{dist} = X_{dist\_norm} Y_{view} \\
  Z_{dist} = Z_{dist\_norm} Y_{view} \\
  \end{gathered}
\end{gather}
```
